---
title: "AI Training Gaps in European Schools"
author: "Ana Nenadic"
format: 
  html:
    embed-resources: true
---

```{r}
#| label: setup
#| message: false
#| warning: false

library(tidyverse)
```

My data is from the OECD [Teaching and Learning International Survey (TALIS) 2024](https://www.oecd.org/en/publications/results-from-talis-2024_90df6235-en/full-report/list-of-tables-available-online_d5b208db.html). Specifically, training gap data is from table 4.30; barriers to AI use is from table 1.63 for middle school teachers across a selection of 28 European countries, further stratified by EU member, EU candidate, and non-member or candidate. 

```{r}
#| label: graph1
#| message: false
#| warning: false

gap <- read_csv("data/training_gap.csv")

#My EU Color Palette
darkblue <- "#3D426B" #for EU member states
lightblue <- "#779ECB" #for Candidate states
grey <- "#DBD9DA"# for non-EU and non-Candidate states
black <- "#000000" #. for general text labels

ggplot(gap, aes (x = reorder(country, training_gap_pct),
                 y = training_gap_pct, fill = group)) +
      geom_col(width=0.8, color = "white", linewidth = 0.3) +
      geom_text (aes(label = paste0(training_gap_pct, "%")),
                 hjust = -0.15, size = 2.6, color = black) +
      coord_flip() +
      scale_fill_manual(values = c("EU member" = darkblue,
                                   "Candidate" = lightblue,
                                   "Other Europe" = grey)) +
      scale_y_continuous(limits = c(0,32)) +
      labs(title = "AI Training Gap in European Schools",
           subtitle = "Middle-School Teachers Reporting Unmet Need For AI Training in 2024",
           x = NULL, y = "% of Teachers",
           fill = "Country group",
           caption = "Source: OECD, Teaching and Learning International Survey 2024") +
      
      theme_minimal()
```

Some Southern European countries have the biggest gaps between need and presence of training: Malta (28.3%), Cyprus (26.9%), and Spain (24.2%). Estonia (9.1%) and several Central European countries like Slovenia (9.4%) and the Czech Republic (10.4%) have the smallest gaps. Meanwhile, EU membership does not delineate the countries since there is a wide range among member states (from 9.1% to 28.3%). Western Balkan candidate countries like Kosovo (21.3%) and North Macedonia (13.4%) are also within that spectrum. This chart is helpful in understanding that EU membership does not necessarily predict how well teacher needs for AI training are met. 
  
```{r}
#| label: graph2
#| message: false
#| warning: false

exp_gap <- read_csv("data/experience_gap.csv")

by_group <- exp_gap |>
  group_by(group) |>
  summarise(le5 = mean(le5_pct), y6_10 = mean(y6_10_pct), gt10 = mean(gt10_pct)) |>
  pivot_longer(cols = c(le5, y6_10,gt10),
               names_to = "experience", values_to ="avg_pct") |>
  mutate(experience = factor(experience, levels = c("le5", "y6_10", "gt10"),
                             labels = c("5 years or less", "6 to 10 years", "More than 10 years")))

ggplot(by_group, aes(x = experience, y = avg_pct, color = group, group = group)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 3) +
  scale_color_manual(values = c("EU member" = darkblue,
                                "Candidate" = lightblue,
                                "Other Europe" = grey)) +
  labs(title = "AI Training Gap Increase With More Experienced Teachers",
       subtitle = "Average Share of Teachers with Unmet AI Training Need by years of Teaching",
       x = "Years of Teaching Experience", y = "% of Teachers",
       color = "Country group",
       caption = "Source: OECD, Teaching and Learning International Survey 2024") +
  
  theme_minimal() +
  theme(legend.position = "bottom")
```
Across the three categories of states in Europe, the share of middle school teachers that report unmet AI training need rises with teaching experience. Specifically, within the EU member state category, those teachers with 5 years or fewer of teaching experience roughly report around 14.4% and those that had more than 10 years report up to 18.9%. Within the EU candidate state category, the difference is similar. On the one end, we have a range from 14.7% to 18.7%. This chart is helpful in understanding that those teachers with most significant experience teaching are, also most likely to report recieving inadequate training; state-wide, and EU-wide, programs can benefit from this information for targeting purposes.
```